\bf Landslide: \\ Systematic Dynamic Race Detection in Kernel Space

### 2012-05-18 conclusion.tex
more copyedit

\revision{
Systematic exploration is a powerful technique for exposing race conditions in concurrent systems.
%It is more reliable than stress testing, and able to find more diverse types of bugs than data race detection.
In previous research, systematic exploration has proven useful in user-level and distributed systems, yet a gap stands between the technique and the complex world of kernel-level concurrency debugging. In this thesis, we present techniques for bridging this gap.

>>>>

\revision{
Systematic exploration is a powerful technique for exposing race conditions in concurrent systems.
%It is more reliable than stress testing, and able to find more diverse types of bugs than data race detection.
In previous research, systematic exploration has proven useful in user-level code and distributed systems, yet a gap stands between the technique and the complex world of kernel-level concurrency debugging. In this thesis, we present techniques for bridging this gap.

### 2012-05-18 evaluation.tex
remove some todos and fixmes
more copyedit

\begin{enumerate}
	\item Annotating their kernel (Section~\ref{sec:using-annotations}).
	\item Implementing instrumentation within Landslide (Section~\ref{sec:using-student-c}).
	\item Fixing problems encountered getting Landslide to run, such as editing their kernel to meet Landslide's requirements and debugging incorrect instrumentation.
	\item Customising Landslide's search parameters such as decision points (Section~\ref{sec:using-customise}).
	\item Analysing the output from bugs that Landslide found.
\end{enumerate}

@@

\newcommand\asdf[1]{\hspace{-0.05in}\footnotesize\bf{#1}\hspace{-0.05in}}
\revision{
\begin{table}[t]
	\begin{center}
	\small
	\begin{tabular}{|l||c|c|c|c|c|c||c|c|}
		\hline
		\multirow{2}{*}{\bf Group} & \multicolumn{6}{|c||}{\bf Minutes spent doing\dots} & \multicolumn{2}{|c|}{\bf Total time spent} \\
		\cline{2-9}
		& \asdf{annotating} & \asdf{config} & \asdf{student.c} & \asdf{fixing} & \asdf{customising} & \asdf{found races} & \asdf{required} & \asdf{refinement} \\
		\hline \hline
		% ntaha/jmchow
		\multirow{2}{*}{Group 1} & 35 & {\footnotesize \em N/A} & {\footnotesize \em N/A} & 5 & {\footnotesize \em N/A} & {\footnotesize \em N/A} & {\footnotesize \em N/A} & {\footnotesize \em N/A} \\
		\cline{2-9}
		& 5 & 30 & 10 & 100 & {\footnotesize \em N/A} & {\footnotesize \em N/A} & 145 & {\footnotesize \em N/A} \\
		\hline
		% echo/margaret
		\multirow{2}{*}{Group 2} & 20 & 30 & 10 & 80 & {\footnotesize \em N/A} & 10 & 140 & 10 \\
		\cline{2-9}
		& 15 & 45 & 30 & 60 & {\footnotesize \em N/A} & 30 & 150 & 30 \\
		\hline
		% tpassaro
		\multirow{2}{*}{Group 3} & 15 & 10 & 5 & 60 & 30 & 30 & 90 & 60 \\
		\cline{2-9}
		& 15 & 10 & 5 & 60 & 30 & 30 & 90 & 60 \\
		\hline
		% pjumde
		\multirow{2}{*}{Group 4} & 10 & 42 & 3 & 5 & 30 & {\footnotesize \em N/A} & 60 & 30 \\
		\cline{2-9}
		& 15 & 35 & {\footnotesize \em N/A} & 5 & 35 & {\footnotesize \em N/A} & {\footnotesize \em N/A} & 35 \\
		\hline
		% apodolsk
		Group 5 & 50 & 15 & {\footnotesize \em N/A} & 15 & {\footnotesize \em N/A} & {\footnotesize \em N/A} & {\footnotesize \em N/A} & {\footnotesize \em N/A} \\
		\hline
		Former TA & 40 & 20 & 3 & 95 & {\footnotesize \em N/A} & 30 & 158 & 30 \\
		\hline
		\hline
		{\bf Average } & {\bf 22 } & {\bf 26.33 } & {\bf 9.43 } & {\bf 48.5 } & {\bf 31.25 } & {\bf 26 } & {\bf 119 } & {\bf 36.43} \\
		\hline
	\end{tabular}
	\end{center}
	\caption{Self-reported times spent by users on each phase while using Landslide. ``N/A'' indicates a student reported spending no time on a particular phase. ``Required'' indicates the sum of the times from the first four phases; ``refinement'' inducates the sum from the last two phases.
	``N/A'' indicates that a user did not report spending any time on a phase; the ``required'' field is ``N/A'' if any of its represented phases are, and ``N/A'' fields are not reflected in the averages.}
	\label{fig:student-times}
\end{table}

@@

% TODO: Investigate the tree size that bugs did exist in. Investigate how many buggy branches there were.

>>>>

\begin{enumerate}
	\item Annotating their kernel (Section~\ref{sec:using-annotations}).
	\item Implementing instrumentation within Landslide (Section~\ref{sec:using-student-c}).
	\item Fixing problems encountered getting Landslide to run, such as revising their kernel to meet Landslide's requirements and debugging incorrect instrumentation.
	\item Customising Landslide's search parameters such as decision points (Section~\ref{sec:using-customise}).
	\item Analysing the output from bugs that Landslide found.
\end{enumerate}

@@

\newcommand\asdf[1]{\hspace{-0.05in}\footnotesize\bf{#1}\hspace{-0.05in}}
\revision{
\begin{table}[t]
	\begin{center}
	\small
	\begin{tabular}{|l||c|c|c|c|c|c||c|c|}
		\hline
		\multirow{2}{*}{\bf Group} & \multicolumn{6}{|c||}{\bf Minutes spent doing\dots} & \multicolumn{2}{|c|}{\bf Total time spent} \\
		\cline{2-9}
		& \asdf{annotating} & \asdf{config} & \asdf{student.c} & \asdf{fixing} & \asdf{customising} & \asdf{found races} & \asdf{required} & \asdf{refinement} \\
		\hline \hline
		% ntaha/jmchow
		\multirow{2}{*}{Group 1} & 35 & {\footnotesize \em --} & {\footnotesize \em --} & 5 & {\footnotesize \em --} & {\footnotesize \em --} & {\footnotesize \em --} & {\footnotesize \em --} \\
		\cline{2-9}
		& 5 & 30 & 10 & 100 & {\footnotesize \em --} & {\footnotesize \em --} & 145 & {\footnotesize \em --} \\
		\hline
		% echo/margaret
		\multirow{2}{*}{Group 2} & 20 & 30 & 10 & 80 & {\footnotesize \em --} & 10 & 140 & 10 \\
		\cline{2-9}
		& 15 & 45 & 30 & 60 & {\footnotesize \em --} & 30 & 150 & 30 \\
		\hline
		% tpassaro
		\multirow{2}{*}{Group 3} & 15 & 10 & 5 & 60 & 30 & 30 & 90 & 60 \\
		\cline{2-9}
		& 15 & 10 & 5 & 60 & 30 & 30 & 90 & 60 \\
		\hline
		% pjumde
		\multirow{2}{*}{Group 4} & 10 & 42 & 3 & 5 & 30 & {\footnotesize \em --} & 60 & 30 \\
		\cline{2-9}
		& 15 & 35 & {\footnotesize \em --} & 5 & 35 & {\footnotesize \em --} & {\footnotesize \em --} & 35 \\
		\hline
		% apodolsk
		Group 5 & 50 & 15 & {\footnotesize \em --} & 15 & {\footnotesize \em --} & {\footnotesize \em --} & {\footnotesize \em --} & {\footnotesize \em --} \\
		\hline
		Former TA & 40 & 20 & 3 & 95 & {\footnotesize \em --} & 30 & 158 & 30 \\
		\hline
		\hline
		{\bf Average } & {\bf 22 } & {\bf 26.33 } & {\bf 9.43 } & {\bf 48.5 } & {\bf 31.25 } & {\bf 26 } & {\bf 119 } & {\bf 36.43} \\
		\hline
	\end{tabular}
	\end{center}
	\caption{Self-reported times spent by users on each phase while using Landslide. ``--'' indicates a student reported spending no time on a particular phase. ``Required'' indicates the sum of the times from the first four phases; ``refinement'' inducates the sum from the last two phases.
	``--'' indicates that a user did not report spending any time on a phase; the ``required'' field is ``--'' if any of its represented phases are, and ``--'' fields are not reflected in the averages.}
	\label{fig:student-times}
\end{table}

@@

### 2012-05-18 futurework.tex
more copyedit

If Landslide becomes integrated into an undergraduate course curriculum, it will be important to minimise the amount of effort students have to go through to instrument their kernels. This could be achieved in part by changing the project definition in ways that make Landslide better able to automatically instrument kernels. For Pebbles, we recommend the following changes.

\begin{enumerate}
	\item {\bf Timer ticks control ``runnable'' threads.} As discussed in Section~\ref{sec:model-timer}. Additionally, to prevent the users from having to instrument non-descheduling mutexes, the project could mandate that all concurrency primitives actually remove blocked threads from the runqueue.
	\item {\bf No idling when progress can be made.} Also as discussed in Section~\ref{sec:model-timer}. With the current specification, inappropriate idling is merely a performance bug, though it is incompatible with Landslide. If the project gave stricter requirements regarding this, fewer students would need to fix problems with their scheduler idling during the instrumentation process.
	\item {\bf Mandated function names for scheduling events.} As part of the instrumentation process, students tell Landslide where in their schedulers notable events happen. Landslide could automatically detect some of these if the project mandated the names of certain functions, such as the timer handler, context switcher, and runqueu manipulation routines.
\end{enumerate}

@@

For example, one of the groups in the user study said they were looking for a race that only arose in out of memory conditions. A test case that could expose that race, by exhausting the kernel's memory resources, would necessarily have a very large state space that would be impractical to explore systematically.

We suggested to the group to configure their kernel to fail memory allocations every so often, so Landslide would be better able to find the race while still using a small test case. To make Landslide more applicable to such bugs in general, we might further investigate this debugging strategy in a more structured way.
}

@@

Compared to executing the code being tested on real hardware (which tools such as \cite{dbug-ssv} are able to do for user-space programs), Simics is {\em slow}, and even slower with Landslide hooked up to it.

>>>>

If Landslide becomes integrated into an undergraduate course curriculum, it will be important to reduce the amount of effort students must go through to instrument their kernels. This could be achieved in part by changing the project definition in ways that make Landslide better able to automatically instrument kernels. For Pebbles, we recommend the following changes.

\begin{enumerate}
	\item {\bf Timer ticks control ``runnable'' threads.} As discussed in Section~\ref{sec:model-timer}. Additionally, to prevent the users from having to instrument non-descheduling mutexes, the project could mandate that all concurrency primitives actually remove blocked threads from the runqueue.
	\item {\bf No idling when progress can be made.} Also as discussed in Section~\ref{sec:model-timer}. With the current specification, inappropriate idling is merely a performance bug, though it is incompatible with Landslide. %If the project gave stricter requirements regarding this, fewer students would need to fix problems with their scheduler idling during the instrumentation process.
	\item {\bf Mandated function names for scheduling events.} As part of the instrumentation process, students tell Landslide where in their schedulers notable events happen. Landslide could automatically detect some of these if the project mandated the names of certain functions, such as the timer handler, context switcher, and runqueue manipulation routines.
\end{enumerate}

@@

For example, one of the groups in the user study said they were looking for a race that only arose in out-of-memory conditions. A test case that could expose that race, by exhausting the kernel's memory resources, would necessarily have a very large state space that would be impractical to explore systematically.

We suggested that the group configure their kernel to fail memory allocations every so often, so Landslide would be better able to find the race while still using a small test case. To make Landslide more applicable to such bugs in general, we might further investigate this debugging strategy in a more structured way.
}

@@

Compared to executing the code being tested on real hardware (which tools such as \cite{dbug-ssv} are able to do for user-space programs), simulation in Simics is slow, and even slower with Landslide hooked up to it.
